2024 US Presidential Election Forecast Model

### 2024-11-04 paper/paper.qmd
Update by Doran
-Delete stater files
-Update paper.qmd and paper.pdf
-Upload new clean_data

clean_data <- read_csv("../data/01-raw_data/clean_data.csv")
raw_data <- read_csv("../data/01-raw_data/president_polls.csv")
nyt_data <- read_csv("../data/02-analysis_data/NYT_data.csv")
```

@@

clean_data |>
  select(1:5) |>      # Select the first 3 columns
  head(5) |>
  kable(
    col.names = names(clean_data)[1:5],  # Automatically get the first 3 column names to match data
    booktabs = TRUE
  )
```

ggplot(clean_data, aes(x = answer)) +
  geom_bar(fill = "red", color = "black") +
  labs(x = "Answer", y = "Count") +
  theme_minimal()
```

ggplot(clean_data, aes(x = answer, y = pct, fill = party)) +
  geom_bar(stat = "identity", color = "black") +
  labs( x = "Candidate", y = "Percentage") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) + # Converts y-axis to percentage format
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotates x-axis labels
```

@@

### Model justification
```{r}
#| echo: false
#| warning: false
#| message: false
trump_data <- raw_data %>% filter(candidate_name == "Donald Trump")
trump_model <- lm(pct ~ pollscore + sample_size + numeric_grade + transparency_score, data = trump_data)
summary(trump_model)
```

```{r}
#| echo: false
#| warning: false
#| message: false
harris_data <- raw_data %>% filter(candidate_name == "Kamala Harris")
harris_model <- lm(pct ~ pollscore + sample_size + numeric_grade + transparency_score, data = harris_data)
summary(harris_model)
```

# Results
## Trump
The model estimates the influence of several predictors on the dependent variable. The intercept is estimated at 48.56, indicating the baseline level of the dependent variable when all predictors are held at zero.

@@

# Appendix {1}

# Appendix {2}
Idealized Methodology and Survey for Forecasting the US Presidential Election

@@

1. Are you a registered voter?

Yes

No

2. Which state do you reside in?

3. What is your age group?

18-24

25-34

35-44

45-54

55-64

65+

4. Which racial or ethnic group(s) do you identify with? (Select all that apply)

White/Caucasian

Black/African American

Hispanic/Latino

Asian/Pacific Islander

Native American

5. What is your highest level of education?

Less than high school

High school diploma or GED

Some college, no degree

Associate degree (e.g., AA, AS)

Bachelor’s degree (e.g., BA, BS)

Master’s degree (e.g., MA, MS, MEd)

Professional degree (e.g., JD, MD)

Doctoral degree (e.g., PhD, EdD)

6. What is your gender?

@@

7. Which candidate do you currently intend to vote for in the upcoming presidential election?

@@

8. What is your political affiliation?

@@

9. How confident are you in your choice for the presidential election?

@@

10. What is the most important issue influencing your vote? (Choose the top 3)

@@

11. How often do you follow news and updates about the election?

@@

End of Survey: Thank you for your time and valuable input. Your responses will contribute significantly to our understanding of public opinion regarding the upcoming presidential election.
# References

>>>>

clean_data <- read_csv("../data/01-raw_data/clean_data.csv")
nyt_data <- read_csv("../data/02-analysis_data/NYT_data.csv")
cleaned_data <- read_csv("../data/02-analysis_data/cleaned_data.csv")
model_data <- read_csv("../data/02-analysis_data/cleaned_data.csv")
```

@@

cleaned_data |>
  select(everything()) |>      # Select the first 3 columns
  head(5) |>
  kable(
    col.names = names(cleaned_data)[1:ncol(cleaned_data)],  # Automatically get the first 3 column names to match data
    booktabs = TRUE
  )
```

@@

### Model justification
```{r}
#| echo: false
#| warning: false
#| message: false
trump_data <- cleaned_data %>% filter(answer == "Trump")
trump_model <- lm(pct ~ pollscore + sample_size + numeric_grade + transparency_score, data = trump_data)
summary(trump_model)
```

```{r}
#| echo: false
#| warning: false
#| message: false
harris_data <- model_data %>% filter(answer == "Harris")
harris_model <- lm(pct ~ pollscore + sample_size + numeric_grade + transparency_score, data = harris_data)
summary(harris_model)
```

# Results

ggplot(cleaned_data %>% filter(answer %in% c("Trump", "Harris")), aes(x = answer)) +
  geom_bar(fill = "black", color = "black") +
  labs(x = "Answer", y = "Count") +
  theme_minimal()
```

ggplot(cleaned_data %>% filter(answer %in% c("Trump", "Harris")), aes(x = answer, y = pct, fill = party)) +
  geom_bar(stat = "identity", color = "black") +
  labs( x = "Candidate", y = "Percentage") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) + # Converts y-axis to percentage format
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotates x-axis labels
```

```{r}
#| label: fig-pct_overtime
#| fig-cap: Mean Percentage of Support for Trump and Harris over time
#| echo: false
#| warning: false
#| message: false
#| 
trump_data <- cleaned_data %>%
  filter(answer == "Trump") %>%
  mutate(end_date = mdy(end_date)) %>%  # Convert end_date to Date format
  group_by(end_date) %>%
  summarize(mean_pct = mean(pct, na.rm = TRUE))

ggplot(trump_data, aes(x = end_date, y = mean_pct)) +
  geom_smooth(method = "loess", color = "red", fill = "#FF7F7F", linetype = "solid", size = 1.3, se = TRUE, ) +
  annotate("text", 
           x = max(trump_data$end_date), 
           y = trump_data$mean_pct[which.max(trump_data$end_date)], 
           label = paste0("Trump ", round(trump_data$mean_pct[which.max(trump_data$end_date)], 1), "%"),
           color = "red", 
           fontface = "bold",
           size = 2) +
  labs(title = NULL, x = "Date", y = "Percentage") +
  theme_minimal() +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

```

## Trump
The model estimates the influence of several predictors on the dependent variable. The intercept is estimated at 48.56, indicating the baseline level of the dependent variable when all predictors are held at zero.

@@

\newpage

\appendix

# Appendix {-}

# Discussion of the survey’s methodology of Siena/NYT

## Population, Frame, and Sample

-Population:

-Frame:

-Sample:
a topline sample size of at least 100

## Sample Recruitment Method

## Sampling Approach's trade-offs

-Strengths:

-Weaknesses:

## How is non-response handled

## What is good and bad about the questionnaire

### Advantages

### Disadvantages

# The idealized Methodology and Survey of forecasting the US presidential election
Idealized Methodology and Survey for Forecasting the US Presidential Election

@@

  Q1. Are you a registered voter?

 \hspace{1cm} Yes

 \hspace{1cm} No

  Q2. Which state do you reside in?

  Q3. What is your age group?

  \hspace{1cm} 18-24

  \hspace{1cm} 25-34

  \hspace{1cm} 35-44

  \hspace{1cm} 45-54

  \hspace{1cm} 55-64

  \hspace{1cm} 65+

  Q4. Which racial or ethnic group(s) do you identify with? (Select all that apply)

  \hspace{1cm}  White/Caucasian

  \hspace{1cm}  Black/African American

  \hspace{1cm}  Hispanic/Latino

  \hspace{1cm}  Asian/Pacific Islander

  \hspace{1cm}  Native American

  \hspace{1cm}  Other (please specify)

  \hspace{1cm}  Prefer not to say

  Q5. What is your highest level of education?

  \hspace{1cm}  Less than high school

  \hspace{1cm}  High school diploma or GED

  \hspace{1cm}  Some college, no degree

  \hspace{1cm}  Associate degree (e.g., AA, AS)

  \hspace{1cm}  Bachelor’s degree (e.g., BA, BS)

  \hspace{1cm}  Master’s degree (e.g., MA, MS, MEd)

  \hspace{1cm}  Professional degree (e.g., JD, MD)

  \hspace{1cm}  Doctoral degree (e.g., PhD, EdD)

  \hspace{1cm}  Prefer not to say

  Q6. What is your gender?

@@

 Q7. Which candidate do you currently intend to vote for in the upcoming presidential election?

@@

  Q8. What is your political affiliation?

@@

  Q9. How confident are you in your choice for the presidential election?

@@

  Q10. What is the most important issue influencing your vote? (Choose the top 3)

@@

  Q11. How often do you follow news and updates about the election?

@@

End of Survey: Thank you for your time and valuable input. Your responses will contribute significantly to our understanding of public opinion regarding the upcoming presidential election.

\newpage

# References
